---
title: "RedWineQuality"
output: html_document
---
========================================================
## Instroduction
In this project, we are going to investigate which chemcial properties impact the most
the quality of red wines.

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
library(ggplot2)
library(dplyr)
library(gridExtra)
library(GGally)
library(memisc)
library(pander)
library(corrplot)
```

# Load data ---------------------------
## Load the Data
```{r echo=FALSE, Load_the_Data}
redwine <- read.csv('wineQualityReds.csv')
head(redwine)
```

## Overview of the data
### list all the variables in the data
```{r echo=FALSE, variables}
names(redwine)
```

### list the structure of the data
List the structure the data to understand the data type for each variable.
Get the number of occurence for each quality grade.

```{r echo=FALSE, data_structure}
str(redwine)
table(redwine$quality)
```

### remove useless index variable X and creating a new variable 'qualityfactor',then verify
As variable X is only a index, I delete it from the dataset in order to accelerate the process speed in the furture analysis.
I created a new factor variable as 'qualityfacotr' for the future plot purpose.
```{r echo=FALSE, data_type_transform}
rw <- subset(redwine, select = -X)
head(rw)
rw$qualityfactor <- factor(rw$quality)
str(rw$qualityfactor)
names(rw)
```


# Plot data ---------------------------
## Univariate Plots Section
### Plot a bar chart of Quality distribution
```{r echo=FALSE, Quality_distribution}
ggplot(data = rw, aes(x = quality, fill = 'red')) +
  geom_bar() +
  xlab('Red wine quality grade') + 
  ylab('Count') +
  ggtitle('Red Wine Quality Distribution')
```

As shown in the chart, the numbers of each quality grade follows a normal 
distribution.
The red wine with quality grade 5 and 6 represent around 80% of the sample, 
while the number of worse and excellent grades are very small.


### Distribution of main chemical properties
#### Plot histogram charts of Different acids distributions
```{r echo=FALSE, Different_acids}
p1<- ggplot(aes(x = fixed.acidity), data = rw) + 
  geom_histogram(fill='darkgreen', binwidth = 0.5) +
  labs(x = "fixed acidity - g/dm^3")

p2<- ggplot(aes(x = volatile.acidity), data = rw) +
  geom_histogram(fill="green", binwidth = 0.05) +
  labs(x = "volatile acidity - g/dm^3")
        
p3<- ggplot(aes(x = citric.acid), data = rw) + 
  geom_histogram(fill="lightgreen", binwidth = 0.02) + 
  labs(x = "citric acid - g/dm^3")

grid.arrange(p1,p2,p3, ncol=3, top = 'Red wine distribution of properties: 
             acids - g/dm^3') 

summary(rw$citric.acid)
```

Both fixed acidity and volatile acidity properties almost follow the normal 
distribution, while citric acid is a bit right skewed with a mode at 0 and a 
median at 0.26, which means the majority of the wine sample possesses very few 
citric acids. 

#### Plot histogram charts of Residual sugar and chlorides distributions
```{r echo=FALSE, residual_sugar_and_chlorides}
p4<- ggplot(aes(x = residual.sugar), data = rw) + 
  geom_histogram(fill="pink", binwidth = 0.1) +
  xlim(0, quantile(rw$residual.sugar, .95)) +
  labs(x = "residual sugar - g/dm^3")

p5<- ggplot(aes(x = chlorides), data = rw) + 
  geom_histogram(fill="blue", binwidth = 0.001) + 
  xlim(0, quantile(rw$chlorides, .95)) +
  labs(x = "chlorides - g/dm^3")

grid.arrange(p4,p5, ncol=2, 
             top = 'Red wine distribution of different chemical properties: 
             residual sugar and chlorides - g/dm^3')
```

Even though the sample for chlorides follows more to Normal distribution, the 
distribution form for both residual sugar and chlorides are similar: long-tailed
with some outliers. I chose to remove these outliers with a quantile of 95% to 
obtain an approximately normal distribution.


#### Plot histogram chart of Dioxide distribution
```{r echo=FALSE, Dioxide}
p6<- ggplot(aes(x = free.sulfur.dioxide), data = rw) + 
  geom_histogram(fill="grey", binwidth = 2) + 
  xlab("free sulfur dioxide - mg/dm^3")

p7<- ggplot(aes(x = total.sulfur.dioxide), data = rw) +
  geom_histogram(fill="#004074", binwidth = 5) + 
  xlim(0, quantile(rw$total.sulfur.dioxide, .98)) +
  xlab("total sulfur dioxide - mg/dm^3")

grid.arrange(p6,p7, ncol=2, 
             top = 'Red wine distribution of different chemical properties: 
             dioxide - mg/dm^3')
```

Today, sulfur dioxide is widely accepted as a useful winemaking aide for 
preservation purpose.
We notice here the sample distributions for free and total sulfur dioxide are 
very similar, being right-skewed. I am going to transform the data using log 10 
to see if the distribution will trend to normal.


#### Adjust dioxide distribution by Log transformation
```{r echo=FALSE, log_transformation_dioxide}
p6_log<- ggplot(aes(x = free.sulfur.dioxide),  data = rw) + 
  geom_histogram(fill="grey") + 
  labs(x = "log(free sulfur dioxide - mg/dm^3)") + 
  scale_x_log10()

p7_log<- ggplot(aes(x = total.sulfur.dioxide), data = rw) + 
  geom_histogram(fill="#004074") + 
  labs(x = "log(total sulfur dioxide - mg/dm^3)") + 
  scale_x_log10()

grid.arrange(p6_log, p7_log, ncol=2, top = 'Red wine distribution of different 
             chemical properties_log(dioxide - mg/dm^3)')
```

After log transformation, the sample distribution for total sulfur dioxide 
looks much better, while it always looks strange for free sulfur dioxide. 

#### Plot histogram charts of other chemical properties distribution
```{r echo=FALSE, density_pH_sulphates_and_alcohol}
p8<- ggplot(aes(x = density), data = rw) + 
  geom_histogram(fill="purple", binwidth = 0.0001) +
  labs(x = 'density - g/cm^3')

p9<- ggplot(aes(x = pH), data = rw) +
  geom_histogram(fill="#993366", binwidth = 0.01) +
  labs(x = 'pH')

p10<- ggplot(aes(x = sulphates), data = rw) +
  geom_histogram(fill="#654477", binwidth = 0.01) + 
  xlim(0.25, quantile(rw$sulphates, .95)) +
  labs(x = 'sulphates - g/dm^3')

p11<- ggplot(aes(x = alcohol), data = rw) +
  geom_histogram(fill="#722717", binwidth = 0.2) +
  labs(x = 'alcohol - % by volume')

grid.arrange(p8,p9,p10,p11, ncol=2, top = 'Red wine distribution of different 
             chemical properties_density, pH, sulphates and alcohol')
```

According to the charts, the density of the red wine sample is between 0.99 to 
1.005 and most of them is around 0.996 to 0.998, which is very close to water.
Typical pH ranges for red wine are between 3.1 and 3.5.
The majority of red wines sample in this dataset have a potassim sulphate level 
around 0.6g/dm^3.
The alcohol percentage varies between approximately 9% and 14%.

Here, the distributions for the other properties seem approximately to a normal 
distribution, while right skewed and long tail for alcohol, so in order to 
transfor the data distribution being toward more normmally, we are going to use 
log transformation.

#### Transformation_alcohol
```{r echo=FALSE, log_Transformation_alcohol}
p11_log <- ggplot(aes(x = alcohol), data = rw) +
  geom_histogram(fill="#722717") +
  labs(x = 'log(alcohol)') +
  scale_x_log10(breaks = seq(min(rw$alcohol, max(rw$alcohol), 0.1)))

p11_sqrt <- ggplot(aes(x = alcohol), data = rw) +
  geom_histogram(fill="#722717") +
  labs(x = 'sqrt(alcohol)') +
  scale_x_sqrt()

p11_log
p11_sqrt
```

Neither transformation on log10 nor squart seem to help the distribution.

## Univariate Analysis
### What is the structure of your dataset?
This dataset contains 1,599 red wines with 11 variables on the chemical 
properties of the wine.  
From the chart, more than 80% of wine is rated between 5 and 6, which are 
categorised as normal quality. 
We can also observe that there are few best and worst samples, the minimum and 
maximum quality rates are 3 and 8.

### What is/are the main feature(s) of interest in your dataset?
### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

I think only distibutions do not help a lot identify the main features, 
I will more or less investigate in all variables.

### Did you create any new variables from existing variables in the dataset?
I created a new variable called 'qualityfactor' for the furture analysis purpose
by facet.

### Of the features you investigated, were there any unusual distributions? 
### Did you perform any operations on the data to tidy, adjust, or change the \
form of the data? If so, why did you do this?

Yes, please refer above.


## Bivariate Plots Section
### Plot boxplots for diffeernt acids by quality
```{r echo=FALSE, Quality_vs_acids}
ggplot(data = rw, aes(x=qualityfactor, y=fixed.acidity)) +
  geom_jitter(alpha = 0.2) +
  geom_boxplot(alpha = 0.1, color = 'blue') + 
  stat_summary(fun.y = mean, geom = 'point', shape = 23, color = 'red', 
               fill = 'black') +
  labs(x = 'quality', 
       y = 'fixed acidity - g/dm^3', 
       title = 'Quality vs fixed acidity')

ggplot(data = rw, aes(x=qualityfactor, y=volatile.acidity)) +
  geom_jitter(alpha = 0.2) +
  geom_boxplot(alpha = 0.1, color = 'blue') + 
  stat_summary(fun.y = mean, geom = 'point', shape = 23, color = 'red', 
               fill = 'black') +
  labs(x= 'quality',
       y = 'volatile acidity - g/dm^3', 
       title = 'Quality vs volatile acidity')

ggplot(data = rw, aes(x=qualityfactor, y=citric.acid)) +
  geom_jitter(alpha = 0.2) +
  geom_boxplot(alpha = 0.1, color = 'blue') + 
  stat_summary(fun.y = mean, geom = 'point', shape = 23, color = 'red', 
               fill = 'black') +
  labs(x = 'quality',
       y = 'citric acid - g/dm^3',
       title = 'Quality vs citric acid')
```

It seems that the fixed acidity is not one of the main features affecting the 
quality of the red wine, the mean of acidity looks similar between different 
quality grades, even there is a small positive effect between them.
However, according to the chart, the quality grade increases when volatile 
acidity level decreases, even the difference becomes smaller between higher 
quality wines. 
While the relationship between volatile acidity and quality is negative, citric
acid has a positive impact on quality: better wines have a higher concentration 
of citric acid.

### Plot boxplots for residual sugar and chlorides by quality
```{r echo=FALSE, Quality_vs_residual_sugar_and_chlorides}
ggplot(data = rw, aes(x=qualityfactor, y=residual.sugar)) + 
  geom_jitter(alpha = 0.2) +
  geom_boxplot(alpha = 0.1, color = 'blue') + 
  stat_summary(fun.y = mean, geom = 'point', shape = 23, color = 'red', 
               fill = 'black') +
  ylim(0.5, quantile(rw$residual.sugar, .95)) + 
  labs(x =  'quality',
       y = 'residual sugar - g/dm^3',
       title = 'Quality vs residual sugar')

ggplot(data = rw, aes(x=qualityfactor, y=chlorides)) + 
  geom_jitter(alpha = 0.2) +
  geom_boxplot(alpha = 0.1, color = 'blue') + 
  stat_summary(fun.y = mean, geom = 'point', shape = 23, color = 'red', 
               fill = 'black') +
  ylim(0.05, quantile(rw$chlorides, .95)) +
  labs(x =  'quality',
      y = 'chlorides - g/dm^3',
      title = 'Quality vs chlorides')
```

From the chart, the residual sugar and chlorides should not be key features for 
quality, as we can see whatever quality is, the residual sugar and the chlorides 
for majority of red wine sample are respectively around 2.5 and 0.08.

### Plot boxplots for different dioxides by quality
```{r echo=FALSE, Quality_vs_free_sulfur_dioxide_and_total_sulfur_dioxide}
ggplot(data = rw, aes(x=qualityfactor, y=free.sulfur.dioxide)) + 
  geom_jitter(alpha = 0.2) +
  geom_boxplot(alpha = 0.1, color = 'blue') + 
  stat_summary(fun.y = mean, geom = 'point', shape = 23, 
               color = 'black', fill = 'red') +
  scale_y_log10() +
  labs(x =  'quality',
      y = 'log(free sulfur dioxide - mg/dm^3)', 
      title = 'Quality vs free sulfur dioxide')

ggplot(data = rw, aes(x=qualityfactor, y=total.sulfur.dioxide)) + 
  geom_jitter(alpha = 0.2) +
  geom_boxplot(alpha = 0.1, color = 'blue') + 
  stat_summary(fun.y = mean, geom = 'point', shape = 23, 
               color = 'black', fill = 'red') +
  scale_y_log10() +
  labs(x = 'quality',
       y = 'log(total sulfur dioxide - mg/dm^3)',
       title = 'Quality vs total sulfur dioxide')
```

Both free and total sulfur dioxide concentration levels do not seem like 
important features for quality, as we can realise on the chart, red wines with 
quality grade at 4 and 8 almost have same dioxide level.
However, we also found that the wines with quality grade 5 possess the highest 
level of dioxide.

### Plot a boxplot for density by quality
```{r echo=FALSE, Quality_vs_density}
ggplot(data = rw, aes(x=qualityfactor, y=density)) + 
  geom_jitter(alpha = 0.2) +
  geom_boxplot(alpha = 0.1, color = 'blue') + 
  stat_summary(fun.y = mean, geom = 'point', shape = 23, color = 'red', 
               fill = 'black') +
  labs(x = 'quality',
       y = 'denstiy - g/cm^3',
       title = 'Quality vs density')
```

Even the density gap between quality grades is not such considerable, we can 
still see that there is a weak negative correlation between them. 
Better quality wines have a smaller density: 0.995 for rate 8 compared with
0.9975 for rate 3.

### Plot a boxplot for pH by quality
```{r echo=FALSE, Quality_vs_pH}
ggplot(data = rw, aes(x=qualityfactor, y=pH)) + 
  geom_jitter(alpha = 0.2) +
  geom_boxplot(alpha = 0.1, color = 'blue') + 
  stat_summary(fun.y = mean, geom = 'point', shape = 23, color = 'red', 
               fill = 'black') +
  labs(x = 'quality', 
       y = 'pH',
       title = 'Quality vs pH') 
```

Even less pH is, better rated the wine was obtained, pH seems not a key element
for the quality, because the pH gap is really very small between bad wines and 
good wines.


### Plot a boxplot for sulphates by quality
```{r echo=FALSE, Quality_vs_sulphates}
ggplot(data = rw, aes(x=qualityfactor, y=sulphates)) + 
  geom_jitter(alpha = 0.2) +
  geom_boxplot(alpha = 0.1, color = 'blue') + 
  stat_summary(fun.y = mean, geom = 'point', shape = 23, color = 'red', 
               fill = 'black') +
  labs(x = 'quality', 
       y = 'sulphates - g/dm^3',
       title = 'Quality vs sulphates')
```

Sulphate is a feature weakly affecting the quality of the red wine, better rated 
red wines contain more sulphates than badly rated ones. 
We also realise a small sulphates quantity difference may change the red wine 
quality grade.


### Plot a boxplot for alcohol by quality
```{r echo=FALSE, Quality_vs_alcohol}
ggplot(data = rw, aes(x=qualityfactor, y=alcohol)) +
  geom_jitter(alpha = 0.2) +
  geom_boxplot(alpha = 0.1, color = 'blue') + 
  stat_summary(fun.y = mean, geom = 'point', shape = 23, color = 'red', 
               fill = 'black') +
  labs(x = 'quality', 
       y = 'alcohol - % by volume',
       title = 'Quality vs alcohol')
```

It seems that except for the bad-rated wine (rating 3 and 4), more alcohol 
contained in the red wine, a better quality grade is obtained.


### Calculate the correlation coefficients betweeen quality and different \
features
Verify our findings with the correlation coefficients for features with normal 
distribution.
```{r echo=FALSE, correlation_coefficients_betweeen_quality_and_features}
cor.test(x = rw$quality, y = rw$fixed.acidity, alternative = 'two.sided', 
         method = 'pearson', conf.level = 0.95)
cor.test(x = rw$quality, y = rw$volatile.acidity, alternative = 'two.sided', 
         method = 'pearson', conf.level = 0.95)
cor.test(x = rw$quality, y = rw$citric.acid, alternative = 'two.sided', 
         method = 'pearson', conf.level = 0.95)
cor.test(x = rw$quality, y = rw$residual.sugar, alternative = 'two.sided', 
         method = 'pearson', conf.level = 0.95)
cor.test(x = rw$quality, y = rw$chlorides, alternative = 'two.sided', 
         method = 'pearson', conf.level = 0.95)
cor.test(x = rw$quality, y = rw$free.sulfur.dioxide, alternative = 'two.sided', 
         method = 'pearson', conf.level = 0.95)
cor.test(x = rw$quality, y = rw$total.sulfur.dioxide, alternative = 'two.sided', 
         method = 'pearson', conf.level = 0.95)
cor.test(x = rw$quality, y = rw$density, alternative = 'two.sided', 
         method = 'pearson', conf.level = 0.95)
cor.test(x = rw$quality, y = rw$pH, alternative = 'two.sided', 
         method = 'pearson', conf.level = 0.95)
cor.test(x = rw$quality, y = rw$sulphates, alternative = 'two.sided', 
         method = 'pearson', conf.level = 0.95)
cor.test(x = rw$quality, y = rw$alcohol, alternative = 'two.sided', 
         method = 'pearson', conf.level = 0.95)
```

Here, the Pearson test confirmed the following features have impacts on 
the quality of red wine: \
- moderate and positive: alcohol \
- moderate and negative: volatile acidity \
- weak and positive: citric acid, sulphates \
- weak and negative: total sulfur dioxide, density \
We also realise the p_value for residual sugar is 0.5832, which is much superior
than alpha = 0.05, so we will not consider its coefficient is statistically 
significant.


### Relationship between features
#### Calculate the correlation coefficients between different features
```{r echo=FALSE, Correlation_coefficients_between_features}
C <- cor(rw[, 1:12])
C
corrplot(C, method = 'circle')
```

From the table above, we found the relationship below:
- fixed acidity: strong positive effect with citric acid and density; 
strong negative with pH, which should be logical, more acides contained, lower 
pH is \
- volatile acidity: strong negative effect with citric acid \
- citric acid: strong negative with pH \
- residual sugar: moderate positive with density, which seems to be normal, 
because sugar will increase the density of the liquid \
- chlorides: moderate positive with sulphates \
- free.sulfur.dioxide: strong positive with total.sulfur.dioxide \
- density: middle negative with alcohol.


#### Plot scatter charts of the realtionship between features

```{r echo=FALSE, acidity_vs_density}
ggplot(data = rw, aes(x=fixed.acidity, y=density)) + 
  geom_jitter(alpha = 0.3) +
  geom_smooth() +
  labs(y = 'denstiy - g/cm^3',
       x = 'fixed acidity - g/dm^3',
       title ='density vs fixed acidity')

ggplot(data = rw, aes(x=volatile.acidity, y=density)) + 
  geom_jitter(alpha = 0.3) +
  geom_smooth() +
  labs(y = 'denstiy - g/cm^3',
       x = 'volatile acidity - g/dm^3',
       title = 'density vs volatile acidity')

ggplot(data = rw, aes(x=citric.acid, y=density)) + 
  geom_jitter(alpha = 0.3) +
  geom_smooth() +
  labs(y = 'density - g/cm^3',
       x = 'citric acid - g/dm^3',
       title = 'density vs citric acid')
```

The fixed acidity has a medium positive effect on the density of the red wine, 
while citric acid has a weak positive relationship with the density.
The volatile acidity seems have no effect on the density.


```{r echo=FALSE, volatile_vs_citric_acid}
ggplot(data = rw, aes(x=volatile.acidity, y=citric.acid)) + 
  geom_jitter(alpha = 0.4) +
  geom_smooth(method = 'lm', color = 'blue') +
  ylim(0, max(rw$citric.acid)) +
  labs(y = 'citric acid - g/dm^3',
      x = 'volatile acidity - g/dm^3', 
      title = 'volatile acidity vs citric acid')
```

The relationship between volatile acidity and citric acid is strongly negative.


```{r echo=FALSE, fixed_acidity_vs_pH}
ggplot(data = rw, aes(x=fixed.acidity, y=pH)) + 
  geom_jitter(alpha = 0.4) +
  geom_smooth(method = 'lm', color = 'pink') +
  labs(x = 'fixed acidity - g/dm^3', 
       y = 'pH',
       title = 'fixed acidity vs pH')
```

The relationship between fixed acidity and pH is medium negative.


```{r echo=FALSE, acidity_vs_alcohol}
ggplot(data = rw, aes(x=fixed.acidity, y=alcohol)) +
  geom_jitter(alpha = 0.4) +
  geom_smooth(color= 'blue') +
  labs(y = 'alcohol - % by volume',
       x = 'fixed acidity - g/dm^3',
       title = 'alcohol vs fixed acidity')

ggplot(data = rw, aes(x=volatile.acidity, y=alcohol)) + 
  geom_jitter(alpha = 0.4) +
  geom_smooth(color= 'blue') +
  labs(y = 'alcohol - % by volume',
       x = 'volatile acidity - g/dm^3',
       title = 'alcohol vs volatile acidity')

ggplot(data = rw, aes(x=citric.acid, y=alcohol)) + 
  geom_jitter(alpha = 0.4) +
  geom_smooth(color= 'blue') +
  labs(y = 'alcohol - % by volume',
       x = 'citric acid - g/dm^3',
       title ='alcohol vs citric acid')
```

It seems that whatever the type of acid, it will not change pH value in the red
wine.


```{r echo=FALSE, alcohol_vs_density}
ggplot(data = rw, aes(x=alcohol, y=density, color= qualityfactor)) + 
  geom_jitter() +
  geom_smooth(method = 'lm', color = 'black') +
  # zoom in by setting ylim
  ylim(0.985, 1.005) +
  scale_color_brewer(type = 'seq') +
  # darken the background for light color
  theme_dark() +
  labs(x = 'alcohol - % by volume',
       y = 'denstiy - g/cm^3',
       title = 'alcohol vs density by colored quality')
```

The red wine with higher alcohol has a lower density.


### Create linear model with some key features
```{r echo=FALSE, linear_model}
m1 <- lm(quality ~ alcohol, data = rw)
m2 <- update(m1, ~ . + fixed.acidity)
m3 <- update(m2, ~ . + volatile.acidity)
m4 <- update(m3, ~ . + citric.acid)
m5 <- update(m4, ~ . + density)
m6 <- update(m5, ~ . + sulphates)
mtable(m1,m2,m3,m4, m5, m6)
```

This linear model is not a good one, since the R2 is only 22.7%, which means 
this model could only explain 22.7% of our red wine dataset.


## Bivariate Analysis
### Talk about some of the relationships you observed in this part of the 
investigation. How did the feature(s) of interest vary with other features in 
the dataset?

We have observed that the volatile acidity has a negative effect on red wine 
quality, while citric acid increases the quality grade. This confirmed that: 
too high level of volatile acidity could lead to an unpleasant taste while small
quantities of citric acid can add 'freshness' and flavour to wines.

Residual sugar and salt in reasonable quantity remained in the wine would not 
affect the quality grade.

pH is neither a key element to the quality because most wines are between 3-4 
on the pH scale.

Wines with lower density and with higher alcohol are rated better quality. We 
also realise that wines with higher alcohol are normally with lower density.

Different acids will not affect alcohol level.

### Did you observe any interesting relationships between the other features 
(not the main feature(s) of interest)?

It is interesting that residual sugar has no significant impact on quality, 
while in the real life, sweety is an important element when we taste the wines. 
If we went deeper, we can see that the residual sugar volume will not change 
alcohol, volatile acidity and other features which impact more or less the red 
wines quality.

### What was the strongest relationship you found?
Alcohol has the strongest positive impact on the quality of the red wines 
compared with other elements, followed by volatile acidity which possess a 
negative relationship with the quality. However, these two elements are 
independent. The volume of volatile acidity would not change a lot the alcohol 
level.


## Multivariate Plots Section
Since alcohol and volatile acidity are two most important elements for quality, 
we are going to investigate them together as well as individually with their the 
most dependent feature.

### Plot a scatter chart of volatile acidity and alcohol by quality grade
```{r echo=FALSE, volatile_acidity_vs_citric_acid}
ggplot(data = rw, aes(x= alcohol, y=volatile.acidity)) + 
  geom_point(aes(color = qualityfactor)) +
  geom_jitter(alpha = 0.1) +
  ggtitle('acidity vs citric acid') +
  scale_color_brewer(type = 'seq') +
  theme_dark() +
  geom_smooth(method = 'lm', color = 'yellow') +
  labs(x = 'alcohol - % by volume',
       y = 'volatile acidity - g/dm^3',
       title = 'relationship between volatile acidity and 
       alcohol by quality grade')
```

Better rated red wines have lower volatile acidity and trend to a higher 
alcohol level. However, the relationship is very weak.


### Plot a scatter chart of volatile acidity and citric acid by quality grade
```{r echo=FALSE, volatile_acidity_vs_citric_acid_by_quality}
ggplot(data = rw, aes(x=volatile.acidity, y=citric.acid)) + 
  geom_point(aes(color = qualityfactor)) +
  geom_jitter(alpha = 0.1) +
   scale_color_brewer(type = 'seq') +
  theme_dark() +
  geom_smooth(method = 'lm', color = 'yellow') +
  ggtitle('volatile acidity vs citric acid') +
  ylim(0,1) +
  labs(x = 'citric acid - g/dm^3',
       y = 'volatile acidity - g/dm^3',
       title = 'relationship between volatile acidity and 
       citric acid by quality') 
```

Better rated red wines contain a higher citric acid level and a lower volatile 
acidity level.


### Plot a scatter chart of density and alcohol by quality grade
```{r echo=FALSE, density_vs_alcohol}
ggplot(data = rw, aes(x=alcohol, y = density)) + 
  geom_point(aes(color = qualityfactor)) + 
   scale_color_brewer(type = 'seq') +
  theme_dark() +
  geom_smooth(method = 'lm', color = 'yellow') +
  labs(x = 'alcohol - % by volume',
       y = 'density - g/cm^3',
       title = 'relationship between density and alcohol by quality') 
```

Better rated red wines contain a higher citric acid level and a lower volatile 
acidity level.

## Multivariate Analysis

### Talk about some of the relationships you observed in this part of the 
investigation. Were there features that strengthened each other in terms of 
looking at your feature(s) of interest?
In the plots above, we can find:
- even the relationship between volatile acidity and density is very weak, 
better rated red wines have lower volatile acidity and trend to a higher 
alcohol level
- better rated red wines contain a higher citric acid level and a lower volatile 
acidity level
- better rated red wines have a lower density and a higher alcohol.

### Were there any interesting or surprising interactions between features?
Normally the density of alcohol is 789kg/m3 and 997kg/m3 for water, which means 
higher alcohol contains, lower density should be. 
So what I found interesting is even the density is not a key feature affecting 
the quality grade, it is a result of the percentage alcohol content of the wine 
and maybe in the next wine quality grade analysis, people no longer need to test
the density.

------

## Final Plots and Summary

### Plot One : plot a histogram chart of red wine quality grade distribution
```{r echo=FALSE, Plot_One}
ggplot(data = rw, aes(x = factor(quality))) +
  geom_bar(color = 'black', fill = 'darkred') +
  labs(x = 'Red wine quality grade: 0_Worst & 10_Best',
       y = 'Count',
       title = 'Red Wine Quality Distribution') +
  geom_text(stat='count', aes(label=..count..), 
            position = position_stack(vjust = 1.15),size=4, color = 'blue')
```

### Description One
The objective of our study is to identify the important features and how they 
affect the quality of the red wine.
Therefore, having an idea of the quality grade distribution of the red wine 
samples is our very first and necessary step.
From the chart, we can see that the distribution almost follows the normal 
distribution.
However, limit to our database volume, the number for badly and 
excellently-rated wines are rare.
80% of the wines are qualified at 5 and 6, which means for a medium quality.

### Plot Two: plot a boxplot of alcool by quality
```{r echo=FALSE, Plot_Two}
ggplot(data = rw, aes(x=qualityfactor, y=alcohol)) +
  geom_jitter(alpha = 0.3) +
  geom_boxplot(alpha = 0.1, color = 'blue') + 
  stat_summary(fun.y = mean, geom = 'point', shape = 23, color = 'red', 
               fill = 'black') +
  labs(y = 'Alcohol level',
      x = 'Red wine quality grade',
      title = 'Quality vs Alcohol by Quality Grade')
```

### Description Two
Our previous analysis confirmed that alcohol is the most important variable in 
the red wine quality grade.
A red wine rated as good (score 7 and 8) contains much higher alcohol than those
who are rated as medium one.
Here, we also find the median and mean alcohol level for bad quality wine seem 
very approximate to the wine rated at 5, but we need to remember the sample 
number for low rated wines are limited and the number for rate 5 is the largest 
(681 out of 1599).

### Plot Three: plot a scatter chart of volatile acidity and citric acid by 
quality grade
```{r echo=FALSE, Plot_Three}
ggplot(data = rw, aes(x=volatile.acidity, y=citric.acid)) + 
  geom_point(aes(color = qualityfactor)) +
  geom_jitter(alpha = 0.1) +
   scale_color_brewer(type = 'seq') +
  theme_dark() +
  geom_smooth(method = 'lm', color = 'yellow') +
  ggtitle('volatile acidity vs citric acid') +
  ylim(0,1) +
  labs(x = 'citric acid - g/dm^3',
       y = 'volatile acidity - g/dm^3',
       title = 'relationship between volatile acidity and 
       citric acid by quality')
```

### Description Three
It is interesting that both being parts of the family 'acids' have totally 
different effects on the quality of the red wine. The volatile acidity has a 
negative effect on the red wine quality, while citric acid increases the quality 
grade.  
High level of volatile acidity could lead to an unpleasant taste while small 
quantities of citric acid can add 'freshness' and flavour to wines.


------

## Reflection
The wine quality is a quite complex analysis and this study is only based on a 
1599 sample. Hence, a linear model could be hard to define exactly how features 
affect the grade. That's why our linear model explains simply 22.7% of data.

On the other hand, a medium quality with 5 and 6 rates representing 80% of the 
sample could also limit the accuracy of our analysis. We do not have enough data
information for low and high-quality wines.

However, we still found alcohol level, volatile acidity, citric acid as well as
sulphates are four relatively more important features to the red wine quality 
grade.

In order to get a more accurate prediction model instead of having a R2 at 22.7%, 
I would suggest collecting a larger database.
Other than that, the majority variables existing in the study seem not key
feature to determine the quality of the red wine, maybe adding more variables
will help, such as grape collection year, region, weather as well as tasters'
information, like gender, age, because people all have different tastes.

### Referrence
- vocabulary: 
chrome-extension://klbibkeccnjlkjkiokjodocebajanakg/suspended.html#ttl=https%3A%
2F%2Fs3.amazonaws.com%2Fudacity-hosted-downloads%2Fud651%2FwineQualityInfo.txt&
uri=https://s3.amazonaws.com/udacity-hosted-downloads/ud651/wineQualityInfo.txt

- alcohol density: https://en.wikipedia.org/wiki/Ethanol
